### 2023-04-05 reachable-resilience.tex
up

\begin{theorem}{(Adjusting Theorem 1 from \cite{DBLP:journals/corr/abs-2108-00889})}\label{post srp}
{\sc state-resilience} is decidable for (not necessarily strong) WSTS with effective 
$\uparrow$ $\post^*$ basis
when
$\Safe=\uparrow \Safe$
and $\Bad=\downarrow \Bad$.
\end{theorem}

@@

Thus in case of a WSTS with an effective basis of $\uparrow \post^*$ and (not strong) compatibility, we don't know the decidability status of $k$-state-resilience and bounded-state-resilience.

\mathieu{Until here}

@@

\begin{theorem}
The state resilience problem is undecidable for effective WSTS with  strong  compatibility 
when
$\Safe=\downarrow \Safe$
and $\Bad=\uparrow \Bad$.
\end{theorem}

\begin{proof}
Consider Reset VASS
in the case $\Safe$ contains only the minimal element %$0$ 
and $\Bad = S \setminus \Safe$, 
it follows the state resilience problem is equivalent 
to the problem of whether $0$ can be visited infinitely often. 
%
% Since reachability of $0$ is undecidable for Reset VASS, we conclude.
Hence the undecidability.
\end{proof}

Despite this, "there is still hope". Indeed, in many ways the case where $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$
is symmetrical to the case $\Safe = \uparrow \Safe$ and $\Bad = \downarrow \Bad$.
%
For instance one can write the following lemma:

@@

\mathieu{Maybe should discuss removing the assumption of upward compatibility and add a sort of 
$\downarrow \post^*$ effectiveness hypothesis, but what would it even mean ?}

\begin{proof}

Procedure 1 enumerates inductive invariants in some fixed order $D_1$ , $D_2$ , . . . , i.e. downward closed subsets $D_i \subseteq S$ such that $\downarrow \post(D_i ) \subseteq D_i$. 
Every inductive invariant $D_i$ is an “over-approximation” of $\downarrow \post^*(s)$ if it contains $s$.
Notice that, by standard monotonicity, $\downarrow \post^*(s)$ is such an inductive invariant and may eventually be found.

@@

Strong downward compatibility implies furthermore the decidability
of $k$-state-resilience and bounded-state-resilience.

\iffalse	\mathieu{}

 The result from \cite{DBLP:journals/corr/abs-2108-00889}
only concerns with the case $\Safe = \uparrow \Safe$
and $\Bad =\downarrow \Bad$, 
 however as mentioned in Section~3
the property for instance of mutual exclusion is often modelised with $\Safe = \downarrow \Safe$.
Hence it makes sense to consider this symmetrical case as well. 
 \cite{DBLP:conf/gg/Ozkan22} considers 
 the case $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$ however the results are ...

\begin{definition}[marked WSTS] 
A {\em marked} WSTS is a tuple $(S, \leq, \to, s_0)$
where $(S, \leq, \to)$ is a WSTS and $s_0 \in S$. 
% If $INIT$ is finite, we call it {\em fin-marked}.
\end{definition}

\begin{definition}
A marked WSTS is
\begin{itemize}
\item lossy if $\downarrow \post^* (s_0) = \post^* (s_0)$,
\item $\bot$-bounded (bottom-bounded) if there is a natural number 
$\ell \geq 0$ s.t. $s_B \in \post^{\leq \ell} (s) $ for every $ s \in S$ and every element 
$s_B$ of a basis of $S$ with $s \geq s_B$.
\end{itemize}
\end{definition}

\begin{theorem}{\cite{DBLP:conf/gg/Ozkan22}}
The bounded-state-resilience problem is decidable for marked
strongly compatible WSTS
 which are 
 {\em \textcolor{red}{lossy} } 
 and 
 {\em \textcolor{red}{$\bot$-bounded}  }
 if $\Bad = \uparrow \Bad$ and $\Safe = \downarrow \Safe$.
\end{theorem}

\mathieu{}	\fi

@@

\iffalse
\mathieu{Should try to put all results in a single table}

With only strong compatibility hypothesis:

\begin{center}
\begin{tabular}{ | l | c | c | r |}
\hline   \Safe~\Bad & $\uparrow$~ $\downarrow$~ & $\downarrow$~ $\uparrow$~  \\ \hline
   SRP & Undecidable (Thm~\ref{srp up down}) & Undecidable  \\ \hline
   BSRP & Undecidable (Corollary~\ref{bsrp up down}) &  ??  \\ \hline
      kSRP & Undecidable (Corollary~\ref{bsrp up down}) & ?? \\ \hline
 \end{tabular}
\end{center}

With $\uparrow \post^*$ effectiveness hypothesis:

\begin{center}
\begin{tabular}{ | l | c | c | r |}
\hline   \Safe~\Bad & $\uparrow$~ $\downarrow$~ & $\downarrow$~ $\uparrow$~  \\ \hline
   SRP & Decidable (Thm~\ref{post srp}) & ??  \\ \hline
   BSRP & ?? &  ??  \\ \hline
      kSRP & ?? & ?? \\ \hline
 \end{tabular}
\end{center}

With $\uparrow \post^*$ effectiveness and strong compatibility hypothesis:

\begin{center}
\begin{tabular}{ | l | c | c | r |}
\hline   \Safe~\Bad & $\uparrow$~ $\downarrow$~ & $\downarrow$~ $\uparrow$~  \\ \hline
   SRP & Decidable (Thm~\ref{postcomputable})& ??  \\ \hline
   BSRP & Decidable (Thm~\ref{ref ozkan})&  ??  \\ \hline
      kSRP & Decidable (Thm~\ref{ref ozkan}) & ?? \\ \hline
 \end{tabular}
\end{center}

@@

Results in the case $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$:

\mathieu{by symmetry} for downward compatible WSTS (but not necessarily upward compatible):

@@

Si on ne veut pas s'embêter avec les "downward compatible pas upward compatible" on peut se contenter d'écrire ça plutôt

>>>>

\begin{theorem}{(Adjusting Theorem 1 from \cite{DBLP:journals/corr/abs-2108-00889})}\label{post srp}
{\sc state-resilience} is decidable for 
%(not necessarily strong)
 WSTS with effective 
$\uparrow$ $\post^*$ basis
when
$\Safe=\uparrow \Safe$
and $\Bad=\downarrow \Bad$.
\end{theorem}

@@

Thus in case of a WSTS with an effective basis of $\uparrow \post^*$ and (not strong) compatibility, we don't know the decidability status of {\sc $k$-state-resilience} and 
{\sc bounded-state-resilience}.

@@

\begin{theorem}
{\sc State-resilience} is undecidable for effective WSTS with  strong  compatibility 
when
$\Safe=\downarrow \Safe$
and $\Bad=\uparrow \Bad$.
\end{theorem}

\begin{proof}
Consider Reset VASS
in the case $\Safe$ contains only the minimal element $Min(S) = 0$ %$0$ 
and $\Bad = S \setminus \Safe$, 
it follows the state resilience problem is equivalent 
to the problem of whether $0$ can be visited infinitely often. 
% Since reachability of $0$ is undecidable for Reset VASS, we conclude.
Hence the undecidability.
\end{proof}

Despite this, it is possible to yield positive results. Indeed, in many ways the case where $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$
is symmetrical to the case $\Safe = \uparrow \Safe$ and $\Bad = \downarrow \Bad$.
%
For instance one can write the following lemma:

@@

\begin{proof}
In order to decide whether state-resilience hold, we execute two procedures in parallel,
one looking for a resilience certificate and one looking for a non-resilience certificate.
Procedure 1 enumerates inductive invariants in some fixed order $D_1$ , $D_2$ , $\ldots$ , i.e. downward closed subsets $D_i \subseteq S$ such that $\downarrow \post(D_i ) \subseteq D_i$. 
Every inductive invariant $D_i$ is an “over-approximation” of $\downarrow \post^*(s)$ if it contains $s$.
Notice that, by standard monotonicity, $\downarrow \post^*(s)$ is such an inductive invariant and may eventually be found.

@@

% \mathieu{Maybe should discuss removing the assumption of upward compatibility and add a sort of $\downarrow \post^*$ effectiveness hypothesis, but what would it even mean ?}

Strong downward compatibility implies furthermore the decidability
of {\sc $k$-state-resilience} and {\sc bounded-state-resilience}.

\begin{proposition}
{\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for ideal effective WSTS with strong downward compatibility,
$\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$.
\end{proposition}

@@

@@

\noindent
Results in the case $\Safe = \downarrow \Safe$ and $\Bad = \uparrow \Bad$:
\iffalse
(by symmetry) for downward compatible WSTS (but not necessarily upward compatible):

@@

Si on ne veut pas s'embêter avec les "downward compatible pas upward compatible" on peut se contenter d'écrire ça plutôt
\fi
